Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-06-23 bbbq_article.tex
Show IC50s per haplotype

The definition of a binder differs from \cite{bianchi2017}:
in this study a peptide is labeled an MHC-I binder if, within a haplotype, 
any of its 9-mer peptides have an IC50 in the lowest 2\% of randomly simulated 
peptides (see Supmat for values \richel{Put these values in supmat}), were the original study defined
a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17

@@

Because we use 13 haplotypes per MHC class,
for a significance level of 5\%, 
there is a 14% chance to find one haplotype 
that hints to have selection working upon it.
% > pbinom(q = 1, size = 13, prob = 0.05, lower.tail = FALSE)
% [1] 0.1354239

The chance to find 2 haplotypes that hint to have selection working upon it
is 2\%, which is below our significance level.
% > pbinom(q = 2, size = 13, prob = 0.05, lower.tail = FALSE)
% [1] 0.02450784

@@

To calculate the probability all of our observations happened by chance,
we view our experiment as a two-state Markov chain, as depicted in
figure \ref{fig:markov_chain}. The two nodes depict the
two possible detection state a peptide is in, which is
either undetected $U$, or detected $D$ by the immune system.
The $D$ can be either detection by an MHC-I or MHC-II haplotype.
The edges depict the probability of a state transition by mutation,
for example, $p_{DU}$ denotes the probability that a detected
peptide mutates into an undetected one.
We focus on peptides that are TMH and we assume a mutation that 
renders the TMH non-functional as lethal.

@@

Of the complete system, we test for each of the 2 state transitions (that is,
we do not test for mutations that result in the same state).
For each transition rate, we know the expected and observed number
of transitions. If the observed number of transitions is below the expected
value, we test for finding that (low) number of transitions or less.
If the observed number of transitions is above the expected
value, we test for finding that (high) number of transitions or more.
We use a significance level $\alpha$ of 5\%, because we have no prior
evidence.

As we are testing 2 hypotheses, 
we correct against type I errors (false positives), 
by applying the Holm-Bonferroni correction, 
resulting in \richel{here is just a normal Bonferroni correction, for now}
$\alpha_c = \frac{\alpha}{2}$.
If the probability of observing that few/many 
transition is below $\alpha_c$, we will state 
that this unlikely situation may be caused by natural selection.

@@

\begin{figure}[!htbp]
  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
  Predicted IC50s (in nM) per MHC-I haplotype
  \caption{
    \richel{Used randomly simulated peptides, results are real}
  }
  \label{fig:ic50s_per_haplotype}
\end{figure}

>>>>

The definition of a binder differs from \cite{bianchi2017}:
in this study a peptide is labeled an MHC-I binder if, within a haplotype, 
any of its 9-mer peptides have an IC50 in the lowest 2\% of randomly simulated 
peptides (see table \ref{table:ic50s_per_haplotype} for values), were the original study defined
a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17

@@

% Chance to find
%            this amount of positives (sums to 1)
%              |
%              v
% > dbinom(x = 0, size = 26, prob = 0.05)
% [1] 0.2635201
% > dbinom(x = 1, size = 26, prob = 0.05)
% [1] 0.3606064
% > dbinom(x = 2, size = 26, prob = 0.05)
% [1] 0.2372411
% > dbinom(x = 3, size = 26, prob = 0.05)
% [1] 0.09989098
% > dbinom(x = 4, size = 26, prob = 0.05)
% [1] 0.03023017
% > dbinom(x = 5, size = 26, prob = 0.05)
% [1] 0.00700067
%
% Chance to find more than
%            this amount of positives (excluding the value!)
%              |
%              v
% > pbinom(q = 0, size = 26, prob = 0.05, lower.tail = FALSE)
% [1] 0.7364799
% > pbinom(q = 1, size = 26, prob = 0.05, lower.tail = FALSE)
% [1] 0.3758735
% > pbinom(q = 2, size = 26, prob = 0.05, lower.tail = FALSE)
% [1] 0.1386324
% > pbinom(q = 3, size = 26, prob = 0.05, lower.tail = FALSE)
% [1] 0.0387414
% > pbinom(q = 4, size = 26, prob = 0.05, lower.tail = FALSE)
% [1] 0.008511231

Because we use 13 haplotypes per MHC class,
for a significance level of 5\%,
for both directions (D to U, or U to D)
we expect 5% times 26 equals 1.3 significant findings.

\richel{Hmm, improve}
The chance to find 2 haplotypes that hint to have selection working upon it
is 4\%, which is below our significance level.
% > pbinom(q = 3, size = 26, prob = 0.05, lower.tail = FALSE)
% [1] 0.0387414

Upon finding 3 haplotypes that hint to have selection working upon it,
there is the additional chance that selection is working in different
directions (i.e. in one case there is selection for losing detection,
where in the other case, there is (unexpected) selection for gaining
detection). Therefore, finding 3 haplotypes that hint to have selection working 
upon it in the same direction

@@

To calculate the probability all of our observations happened by chance,
we view our experiment as a two-state Markov chain, as depicted in
figure \ref{fig:markov_chain}. The two nodes depict the
two possible detection state a peptide is in, which is
either undetected $U$, or detected $D$ by the immune system,
where a detected peptide is a binder, that is, 
a peptide with an IC50 in the lowest 2\%.
The $D$ can be either detection by an MHC-I or MHC-II haplotype.
The edges depict the probability of a state transition by mutation,
for example, $p_{DU}$ denotes the probability that a detected
peptide mutates into an undetected one.
We focus on peptides that are TMH and we assume a mutation that 
renders the TMH non-functional as lethal.

@@

Of the complete system, we test for each of the 2 observed 
state transition rates (that is, we do not test for mutations that result in the same state)
if these are likely to be caused by chance.
For each transition rate, we know the expected and observed number
of transitions. If the observed number of transitions is below the expected
value, we test for finding that (low) number of transitions or less.
If the observed number of transitions is above the expected
value, we test for finding that (high) number of transitions or more.
We use a significance level $\alpha$ of 5\%, because we have no prior
expectations.

As we are testing 2 rates (and thus 2 hypotheses), 
we correct against type I errors (false positives), 
by applying the Bonferroni correction, 
resulting in $\alpha_c = \frac{\alpha}{2}$.
If the probability of observing that few/many 
transition is below $\alpha_c$, we will state 
that this unlikely situation may be caused by natural selection.

@@

\begin{figure}[!htbp]
  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
  Predicted IC50s (in nM) per haplotype
  \caption{
    \richel{Used randomly simulated peptides, results are real}
  }
  \label{fig:ic50s_per_haplotype}
\end{figure}

\begin{table}[!htbp]
  \input{ic50s_per_haplotype/ic50s_per_haplotype.latex}
  \caption{
    Predicted IC50s (in nM) per haplotype
    \richel{Used randomly simulated peptides, results are real}
  }
  \label{table:ic50s_per_haplotype}
\end{table}
